---
title: "Cricket Analysis"
author: "Kathirmani Sukumar"
date: "May 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load data
Use read.csv to load any csv file
```{r}
odi = read.csv('e:/datasets/odi-batting.csv')
#View(odi)
class(odi)
colnames(odi)
str(odi)
nrow(odi)
nlevels(odi$Country)
levels(odi$Country)
unique(odi$Country)
```

### Top ten players based on total runs
```{r}
library(dplyr)
library(ggplot2)

top_players = odi %>% group_by(Player) %>% summarise(total_runs=sum(Runs), avg_runs=mean(Runs)) %>% arrange(-total_runs) %>% head(5)
top_players
```

```{r}
odi %>% group_by(Player) %>% summarise(total_runs=sum(Runs), avg_runs=mean(Runs), matches=n()) %>% arrange(-total_runs) %>% head(5)
```


```{r}
colnames(odi)
```


```{r}
odi %>% group_by(Country, Player) %>% summarise(total=n()) %>% arrange(-total) %>% head(5)
```


### Player's Performance
```{r echo=FALSE}
odi$Century = ifelse(odi$Runs>99,1,0)
odi$Ducks = ifelse(odi$Runs==0,1,0)
odi$Fifty = ifelse(odi$Runs>49&odi$Runs<100,1,0)
odi$missed_century = ifelse(odi$Runs>90&odi$Runs<100,1,0)
#View(odi)

player_summary = odi %>% group_by(Player) %>% summarise(Runs=sum(Runs, na.rm=T),
                                       matches=n(),
                                       avg_runs=mean(Runs, na.rm=T),
                                       Century=sum(Century, na.rm=T),
                                       Ducks=sum(Ducks, na.rm=T),
                                       Fifties=sum(Fifty, na.rm=T),
                                       missed_century=sum(missed_century, na.rm=T)
                                       ) %>% arrange(-Runs)
player_summary
#write.csv(player_summary, 'player_summary.csv')
#getwd()
```
- Sachin has most number missed centuries 
- Jayasuriya has most number of ducks

## Plotting
```{r}
library(ggplot2)
top_players = odi %>% group_by(Player) %>% summarise(total_runs=sum(Runs, na.rm=T)) %>% arrange(-total_runs) %>% head(10)

ggplot(top_players, aes(x=reorder(Player, total_runs), y=total_runs)) + geom_bar(stat='identity', fill='steelblue')

players_plt = ggplot(top_players, aes(x=reorder(Player, total_runs), y=total_runs))
players_plt = players_plt + geom_bar(stat='identity', fill='steelblue')
players_plt = players_plt + xlab('Player Name') + ylab('Total Runs') + coord_flip()
plot(players_plt)
```


```{r}
odi %>% group_by(Player) %>% summarise(total_runs=sum(Runs, na.rm=T)) %>% arrange(-total_runs) %>% head(10) %>% ggplot(aes(x=Player, y=total_runs))+ geom_bar(stat='identity')
```

```{r}
ggplot(odi, aes(x=Country))+geom_bar()

```

## Box plot
```{r}
boxplot(odi$Runs)
```

### Histogram
```{r}
hist(odi$Runs)
```
### Identify missing values
```{r}
sum(is.na(odi$Runs))

colSums(is.na(odi)) / nrow(odi) * 100
```

### Categorical column
```{r, fig.width=8, fig.height=4}
odi %>% group_by(Player) %>% summarise(count=n()) %>% head(10) %>% ggplot(aes(x=Player, y=count))+geom_bar(stat='identity', fill='steelblue') + coord_flip()
```

### Pareto Analysis

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(qcc)
pareto.chart(table(odi$Country))
```

## Dates
- increment dates
- diff between dates
- Minimum date that support

```{r}
odi$date = as.Date(odi$MatchDate, "%m-%d-%Y")
head(odi$date)
class(odi$date)
odi$year = format(odi$date, '%Y')
odi$month = format(odi$date, '%m')
odi$month_year = format(odi$date, '%m-%Y')
odi$weekday = format(odi$date, '%A')
head(odi$weekday)
# odi$weekday = weekday(odi$date)
```

### Word cloud
```{r warning=FALSE, echo=FALSE}
library(tm)
library(wordcloud)
library(RColorBrewer)

tweets = read.csv('e:/datasets/narendramodi_tweets.csv')
tweets$text = tolower(tweets$text)
tweets$text = gsub('[^a-zA-Z0-9#& ]', '', tweets$text)
row_words = strsplit(tweets$text, split = ' ')
words = unlist(row_words)
words_freq = table(words)
words_freq = as.data.frame(words_freq)
# colnames(words_freq) = c('WORDS', 'FREQUENCY')

words_freq %>% arrange(-Freq) %>% head(10)
stop_words = stopwords()

words_freq %>% filter(words == 'modi') # Filter based on one element
words_freq %>% filter(words %in% c('modi', 'the', 'india')) # Filter based on more than one element

words_freq = words_freq %>% filter(!words %in% stop_words)
words_freq %>% arrange(-Freq) %>% head(10)
custom_words = c('amp;', 'rt')
words_freq = words_freq %>% filter(!words %in% custom_words)
words_freq %>% arrange(-Freq) %>% head(10)
wordcloud(words_freq$words,
          words_freq$Freq,
          min.freq = 50,
          colors = brewer.pal(8, "Dark2"),
          random.color = T,
          scale = c(10,1)
          )


```

```{r}
library(RSentiment)
Sys.setenv(JAVA_HOME='C:/Program Files/Java/jre-9.0.4') # for 64-bit version
calculate_sentiment(c('I love india. But I hate politics here', 'I hate apple', 'This food is not good'))
calculate_score(c('I love india. But I hate politics here', 'I hate apple', 'This food is not good'))
calculate_sentiment(head(tweets$text))
```


```{r}
library(ggmap)
geocode('RMZ Ecoworld, Bangalore')
geocode('India')
geocode('Bangalore')
uniq_countries = unique(as.character(odi$Country))
geocode(c('India', 'Australia', 'Bangalore'))
countries_geo = geocode(uniq_countries)
countries_geo
write.csv(countries_geo, 'locations.csv')
```

### Bivariate Analysis
- Numerical vs Numerical
- Categorical vs Categorical
- Categorical vs Numerical
```{r}
sachin = odi %>% filter(Player %in% c('Sachin R Tendulkar'))
# sachin = odi %>% filter(Player == 'Sachin R Tendulkar')
nrow(sachin)

ggplot(sachin, aes(x=Runs, y=ScoreRate)) + geom_point(col='steelblue') + geom_smooth(method='lm', se = F)
ggplot(sachin, aes(x=Runs, y=ScoreRate)) + geom_point(col='steelblue') + geom_smooth(method='loess', se = F)

odi %>% filter(Player %in% c('Virender Sehwag')) %>% ggplot(aes(x=Runs, y=ScoreRate))+geom_smooth(se=F)

odi %>% filter(Player %in% c('Virender Sehwag', 'Sachin R Tendulkar', 'Ashish Nehra')) %>% ggplot(aes(x=Runs, y=ScoreRate, col=Player))+geom_smooth(se=F)

odi %>% filter(Player %in% c('Virender Sehwag', 'Sachin R Tendulkar', 'Ashish Nehra')) %>% ggplot(aes(x=Runs, y=ScoreRate, col=Player))+geom_point()
```



```{r}
odi %>% filter(Player %in% c('Virender Sehwag', 'Sachin R Tendulkar', 'Ashish Nehra')) %>% ggplot(aes(x=Runs, y=ScoreRate))+geom_point(col='steelblue' )+facet_wrap(~Player)
```


```{r}
top_players = odi %>% group_by(Player) %>% summarise(centuries=sum(Century, na.rm=T)) %>% arrange(-centuries) %>% head(10)

top_players

odi %>% filter(Player %in% top_players$Player) %>% ggplot(aes(x=Runs, y=ScoreRate))+geom_point(col='steelblue')+facet_wrap(~Player) + geom_smooth()
```

```{r}
tweets %>% ggplot(aes(x=favorite_count, y=retweets_count, ))+geom_point()+facet_wrap(~source)
```

```{r}
boxplot(tweets$favorite_count)
```
### Correlation matrix using corrplot
```{r}
library(corrplot)


stocks = read.csv('e:/datasets/stock-prices.csv')
cor_mat = cor(na.omit(stocks %>% select(-Date)))
cor_mat
```

```{r}
corrplot(cor_mat, method='number')
corrplot(cor_mat, method='color', order = 'hclust', addrect = 2)
```

```{r}
na.omit(stocks)
```

## Categorical vs Numerical
- Boxplot
- Group & averages and plot bar charts
```{r, fig.width=12, fig.height=5}
hr = read.csv('e:/datasets/ibm-watson/HR Analytics.csv')
ggplot(hr, aes(x=MaritalStatus, y=MonthlyIncome))+geom_boxplot()+facet_wrap(~JobRole)
```


```{r, fig.width=12}
sachin %>% group_by(year) %>% summarise(total_runs=sum(Runs, na.rm=T), centuries=sum(Century, na.rm=T)) %>% ggplot(aes(x=year, y=total_runs))+geom_line(group=1, col='steelblue')
```


```{r, fig.width=12}
odi %>% filter(Player %in% head(top_players$Player, 5)) %>% group_by(Player, year) %>% summarise(total_runs=sum(Runs, na.rm=T)) %>% ggplot(aes(x=year, y=total_runs, col=Player))+geom_line(group=1) 
```

## Heatmaps
- Categorical vs Categorical
```{r, fig.width=11}
hr %>% group_by(JobRole, EducationField) %>% summarise(count=n()) %>% ggplot(aes(x=JobRole, y=EducationField, fill=-count))+geom_tile()
```

## Multivariate Analysis
### Scatter Plot
```{r, fig.width=11}
parliament = read.csv('e:/datasets/parliament.csv')

summary_2009 = parliament %>% filter(YEAR==2009, STATE=='KARNATAKA', PC=='BANGALORE CENTRAL')

summary_2009 %>% group_by(NAME) %>% summarise(per_votes=VOTES/sum(summary_2009$VOTES)*100) %>% arrange(-per_votes)

parliament %>% filter(YEAR==2009, Position==1) %>% group_by(PARTY) %>% summarise(wins=n()) %>% arrange(-wins) %>% head(10)

party_wins = parliament %>% filter(Position==1) %>% group_by(YEAR, PARTY) %>% summarise(wins=n())
head(party_wins, 20)


top_parties = party_wins %>% group_by(PARTY) %>% summarise(wins=sum(wins)) %>% arrange(-wins) %>% head(5)
party_wins$YEAR = as.factor(party_wins$YEAR)
ggplot(party_wins %>% filter(PARTY %in% top_parties$PARTY), aes(x=YEAR, y=wins, fill=PARTY))+geom_bar(stat='identity', position = 'stack')
```

```{r, fig.width=11, fig.height=4}
ggplot(party_wins %>% filter(PARTY %in% top_parties$PARTY), aes(x=YEAR, y=wins, fill=PARTY))+geom_bar(stat='identity', position = 'fill')
```

```{r, fig.width=10, fig.height=6}
library(treemap)

players_summary = odi %>% group_by(Player) %>% summarise(total_runs=sum(Runs, na.rm=T), avg_rate=mean(ScoreRate, na.rm=T), Country=first(Country), centuries=sum(Century, na.rm=T)) %>% arrange(-total_runs) %>% head(100)
treemap(players_summary,
        index = c('Country','Player'),
        vSize = 'total_runs',
        vColor = 'centuries',
        type='value'
        )


```

```{r}

```

